---
title: "Riffomanas: The data are?"
author: "Andrew Borozan"
date: "`r Sys.Date()`"
output: html_document
---

I will be following along to Pat Schloss at Riffomanas. [Here is a link to the video](https://www.youtube.com/watch?v=-C8cBbw7tq0&ab_channel=RiffomonasProject). And [here is a link to the web page with the activities.](https://riffomonas.org/code_club/2020-04-02-the-data-are)

```{r Loading Packages}
library(pacman)
p_load(tidyverse, skimr, scales, rmarkdown, magrittr, lubridate, janitor, ggrepel, ggthemes, showtext, knitr, reactable, palmerpenguins, car, broom, ggpubr, readxl, stringr, viridis, gghighlight)
```

How do we get data into a format that we can work with? This is the main question Pat is trying to answer today.

rename() and recode() are going to be the functions we will be practicing today.

Let's load the data we will be using.

```{r Loading the data}
comma <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/comma-survey/comma-survey.csv")

head(comma)

str(comma)
```

The column names for this table are way too long and have capitalization and punctuation which are messy for column names. We need to modify the column names with rename().

rename() allows us to rename columns (seems intuitive).

Let's look at our column names.

```{r Column names}

colnames(comma)
# Gross. 

# For rename() the thing to remember is newname = oldname. Do this for each column you want to change.
comma <- comma %>% 
  rename(respondent = "RespondentID",
         oxford_or_not = "In your opinion, which sentence is more gramatically correct?")
```

recode() allows us to change values within the columns without actually changing the data. That sounds weird but a simple example will hopefully illustrate. Some of the columns have the values "Yes" and "No." If we want those to be "TRUE" and "FALSE" we can use recode() to do so. It doesn't change the value, just the word or input we are using to denote the value.

The syntax is opposite of rename() though. For recode() you do:

-   Column you want to change

-   Old value =

-   New value

(In rename() it is new = old).

```{r}

comma <- comma  %>% 
  mutate(oxford_or_not= 
           dplyr::recode(oxford_or_not, 
                         "It's important for a person to be honest, kind and loyal."="non_oxford",
					             	"It's important for a person to be honest, kind, and loyal."="oxford"))
# Had to put dplyr:: in front of recode because there was another package that has recode and it was using that one. 

comma %>% count(oxford_or_not)
```

On to the exercises. First, rename all of the columns using rename().

```{r}
colnames(comma)

comma <- comma %>% 
  rename(heard_of_oxford = "Prior to reading about it above, had you heard of the serial (or Oxford) comma?",
         care_about_oxford = "How much, if at all, do you care about the use (or lack thereof) of the serial (or Oxford) comma in grammar?",
         write_sentence = "How would you write the following sentence?",
         data_sing_plural = "When faced with using the word \"data\", have you ever spent time considering if the word was a singular or plural noun?",
         care_about_data = "How much, if at all, do you care about the debate over the use of the word \"data\" as a singluar or plural noun?",
         proper_grammar_importance = "In your opinion, how important or unimportant is proper use of grammar?",
         gender = "Gender",
         age = "Age",
         household_income = "Household Income",
         education = "Education",
         location = "Location (Census Region)")

comma %>% colnames()
```

Then, we recode some answers.

```{r}

comma <- comma %>% 
  mutate(care_about_oxford = as.factor(care_about_oxford), 
         write_sentence = 
           dplyr::recode(write_sentence, 
                         "Some experts say it's important to drink milk, but the data are inconclusive." = "data are",
                         "Some experts say it's important to drink milk, but the data is inconclusive." = "data is"),
         care_about_data = as.factor(care_about_data))


```

Extension exercise:

```{r out.width="100%"}

comma %>% drop_na(oxford_or_not) %>% 
  count(oxford_or_not) %>% 
  mutate(percent = percent(n/sum(n)))
```
