---
title: "Cyclistic Project"
author: "Andrew Borozan"
date: "2/10/2022"
output: html_document
---

# Cyclistic Exploration

The following is an exploration project completed at the end of my Google Data Analytics Certification. Here was the scenario:

You are a junior data analyst working in the marketing analyst team at Cyclistic, a bike-share company in Chicago. The director of marketing believes the company's future success depends on maximizing the number of annual memberships. Therefore, your team wants to understand how casual riders and annual members use Cyclistic bikes differently. From these insights, your team will design a new marketing strategy to convert casual riders into annual members. But first, Cyclistic executives must approve your recommendations, so they must be backed up with compelling data insights and professional data visualizations.

The central business task of the project is to increase membership rates.

For the task to be completed, analysis must address the central question: How do annual members and casual riders use Cyclistic bikes differently?

## Setup R Environment

First, we will load the packages that will aid our data exploration. First, load the pacman package, as there is a helpful function in this package that will load other packages needed in the environment.

```{r Load packages, warning=FALSE}
library(pacman)
p_load(tidyverse, skimr, scales, rmarkdown, magrittr, lubridate, 
       janitor, htmltab, ggrepel, viridis, ggthemes, readxl, knitr, showtext,
       showtextdb, reactable, patchwork, reactablefmtr)
```

## Loading Data

The data comes from Cyclistic in the form of .csv files. We load each month into an object and then combine all of the rows into one data frame.

```{r Load in data, include=FALSE}
Cyclistic_January <- read_csv("C:/Users/andre/OneDrive/Desktop/Github/andrewborozanprojects/andrewborozanprojects/Data/Divvy/202101-divvy-tripdata.csv", trim_ws = TRUE)
Cyclistic_February <- read_csv("C:/Users/andre/OneDrive/Desktop/Github/andrewborozanprojects/andrewborozanprojects/Data/Divvy/202102-divvy-tripdata.csv", trim_ws = TRUE)
Cyclistic_March <- read_csv("C:/Users/andre/OneDrive/Desktop/Github/andrewborozanprojects/andrewborozanprojects/Data/Divvy/202103-divvy-tripdata.csv", trim_ws = TRUE)
Cyclistic_April <- read_csv("C:/Users/andre/OneDrive/Desktop/Github/andrewborozanprojects/andrewborozanprojects/Data/Divvy/202104-divvy-tripdata.csv", trim_ws = TRUE)
Cyclistic_May <- read_csv("C:/Users/andre/OneDrive/Desktop/Github/andrewborozanprojects/andrewborozanprojects/Data/Divvy/202105-divvy-tripdata.csv", trim_ws = TRUE)
Cyclistic_June <- read_csv("C:/Users/andre/OneDrive/Desktop/Github/andrewborozanprojects/andrewborozanprojects/Data/Divvy/202106-divvy-tripdata.csv", trim_ws = TRUE)
Cyclistic_July <- read_csv("C:/Users/andre/OneDrive/Desktop/Github/andrewborozanprojects/andrewborozanprojects/Data/Divvy/202107-divvy-tripdata.csv", trim_ws = TRUE)
Cyclistic_August <- read_csv("C:/Users/andre/OneDrive/Desktop/Github/andrewborozanprojects/andrewborozanprojects/Data/Divvy/202108-divvy-tripdata.csv", trim_ws = TRUE)
Cyclistic_September <- read_csv("C:/Users/andre/OneDrive/Desktop/Github/andrewborozanprojects/andrewborozanprojects/Data/Divvy/202109-divvy-tripdata.csv", trim_ws = TRUE)
Cyclistic_October <- read_csv("C:/Users/andre/OneDrive/Desktop/Github/andrewborozanprojects/andrewborozanprojects/Data/Divvy/202110-divvy-tripdata.csv", trim_ws = TRUE)
Cyclistic_November <- read_csv("C:/Users/andre/OneDrive/Desktop/Github/andrewborozanprojects/andrewborozanprojects/Data/Divvy/202111-divvy-tripdata.csv", trim_ws = TRUE)
# write_csv(X202112_divvy_tripdata, file = "C:/Users/andre/OneDrive/Desktop/Github/andrewborozanprojects/andrewborozanprojects/Data/Divvy/202112-divvy-tripdata.csv")
Cyclistic_December <- read_csv("C:/Users/andre/OneDrive/Desktop/Github/andrewborozanprojects/andrewborozanprojects/Data/Divvy/202112-divvy-tripdata.csv", trim_ws = TRUE)


```

```{r Complete Data Set}
Cyclistic <- bind_rows(Cyclistic_January, Cyclistic_February, Cyclistic_March, Cyclistic_April, Cyclistic_May, Cyclistic_June, Cyclistic_July, Cyclistic_August, Cyclistic_September, Cyclistic_October, Cyclistic_November, Cyclistic_December)

head(Cyclistic)

str(Cyclistic)
```

## Cleaning the Data

Next, we clean the data to prepare it for analysis. We will complete the following steps:

-   Remove any duplicates

-   Filter out odd data (e.g. rides that go backwards through time)

-   Remove entries with incomplete data (remove NAs)

-   Create columns that may help further analysis. For example, we will create a ride length column (in four different formats) and extract the month, day, and hour start time.

-   Rearrange columns

-   Filter out any rides that have a ride time of less than 30 seconds (some ride lengths were 1 second long. This is unrealistic. We will set the minimum ride length to 30 seconds.)

```{r Cleaning the data set}
# Remove duplicates
Cyclistic <- Cyclistic %>% distinct(.keep_all = TRUE)

# Filter out any oddities (some rides start after they end somehow)
Cyclistic <- Cyclistic %>% filter(ended_at > started_at)

Cyclistic <- Cyclistic %>% mutate(ride_length = ended_at - started_at,
                                  ride_length_period = seconds_to_period(ride_length), 
                                  ride_length_numeric = as.numeric(ride_length), 
                                  ride_length_char = sprintf('%02d %02d:%02d:%02d', day(ride_length_period), ride_length_period@hour, minute(ride_length_period), second(ride_length_period)),
                                  month = month(started_at), 
                                  day = day(started_at), 
                                  hour = hour(started_at), 
                                  weekday = wday(started_at, label=TRUE, abbr=FALSE)) 


Cyclistic <- Cyclistic %>% select(1:4, month, day, hour, weekday, ride_length, 
                                  ride_length_period, ride_length_numeric, ride_length_char, start_station_name, 
                                  start_station_id, end_station_name, end_station_id, start_lat, start_lng, 
                                  end_lat, end_lng, member_casual) %>% 
  clean_names()
# This rearranges the columns and then cleans the column names

# Drop NAs, and make sure that the ride starts and ends at a different station (some start and end stations are the same)
Cyclistic <- Cyclistic %>% drop_na() %>% filter(ride_length_numeric > 29)


#Skim() from the skimr package will give us a snapshot at some of patterns for each variable. 
skim(Cyclistic)
```

And now, we start looking for some of the behavioral patterns of both members and casual riders.

## Total rides and average length of ride

First, we look at both the total number of rides by members and casual riders, and the mean and median length of ride for each type of rider.

```{r Total Rides and Average Length of Ride by Type of Rider, echo=TRUE}

#Mean and median length of rides by rider type
Cyclistic %>%  
  group_by(member_casual) %>% 
  summarize(count = n(), 
            ride_time_avg = mean(ride_length, na.rm = TRUE),
            ride_time_median = median(ride_length, na.rm = TRUE)) %>%
  mutate(ride_time_avg = round(seconds_to_period(ride_time_avg), digits = 0),
         ride_time_median = round(seconds_to_period(ride_time_median), digits = 0))

```

We see that the mean and median length of rides for casual riders is higher than members. The mean is significantly higher than the median for casual riders. Let's look at the data in a quick box plot to see the spread.

```{r Box Plot for rider type}
options(scipen=999) #Turns off scientific notation for environment

Cyclistic %>% group_by(member_casual) %>% 
  ggplot(aes(member_casual, ride_length_numeric))+
  geom_boxplot() +
  coord_flip()+
  scale_y_log10() + # Have to put in log scale to accomodate range/variability
  theme_classic() +
  xlab(label = "Rider Type") +
  ylab(label = "Ride Length in Seconds")
```

Variability among casual riders is rather high.

## Docked Bikes?

Our data might need further cleaning because of the presence of docked bikes. This problem becomes evident when we take a look at the maximum ride length.

To see the maximum ride length...

```{r Maximum Ride}
maxride <- Cyclistic[which.max(Cyclistic$ride_length), ]

glimpse(maxride)

maxride$ride_length_period
```

That's a really long "ride." The ride type though is "docked bike" meaning this was a bike that was just sitting around for 38 days. If we take a look, docked bikes are only attributed to casual riders, possibly suggesting docked bikes don't get tracked for members.

```{r Number of Docked Bikes by Rider Type, echo=TRUE, message=FALSE, warning=FALSE}
Docked_Bike_Count <- Cyclistic %>% group_by(member_casual, rideable_type) %>% 
  filter(rideable_type == "docked_bike") %>% summarize(count = n())

reactable(Docked_Bike_Count, bordered = TRUE,
  highlight = TRUE)
```

We are going to filter out docked bikes as it seems it will not tell us much about the differences in behavior between members and causal riders, and therefore, gain us little achieving our primary objective.

```{r Filter out Docked Bikes}
Cyclistic <- Cyclistic  %>% 
  filter(rideable_type != "docked_bike")
```

Though we are filtering out docked bikes, this might be a note for marketing strategy. Highlighting the benefit of membership to casual riders who dock frequently and/or for long periods of time might increase rates of membership. All docked ride types are from casual riders (after excluding the one member, which is presumed to be a faulty data point).

Now that we have a data set that has removed incomplete information from members_casual, end_station and start_station ids, and ride_lengths, and we have removed docked bikes, we have a collection of data that can tell us something about the behavior of both members and casual riders. We have already seen that casual riders have much longer ride times on average than members. But that had included docked bikes in the data. What happens when we exclude docked bikes from the data?

```{r Cleaned Mean, Median, and Boxplot}

Cyclistic_CT <- Cyclistic %>%  
  group_by(member_casual) %>% 
  summarize(number_of_rides = n(), ride_time_avg = mean(ride_length, na.rm = TRUE), 
            ride_time_median = median(ride_length, na.rm = TRUE)) %>% 
  mutate(ride_time_avg = round(seconds_to_period(ride_time_avg), digits = 0),
         ride_time_median = round(seconds_to_period(ride_time_median), digits = 0))

Cyclistic_CT


Cyclistic %>% group_by(member_casual) %>% 
  ggplot(aes(member_casual, ride_length_numeric))+
  geom_boxplot() +
  coord_flip()+
  scale_y_log10() + # Have to put in log scale to accomodate range/variability
  theme_classic() +
  xlab(label = "Rider Type") +
  ylab(label = "Ride Length in Seconds")
         

```

We can see a decrease in the amount of rides and average ride time (especially for casual riders), but this data is more robust and includes information we need for analysis.

Visually, the data looks like this:

```{r Total rides by rider type, message=FALSE, warning=FALSE, out.width= "100%"}
font_add_google("Exo", "exo")
showtext_auto()
number <- 
  ggplot(Cyclistic_CT, 
       aes(x = member_casual, 
           y = number_of_rides, fill = member_casual)) + 
  geom_bar(stat = "identity") +
# Make sure scales package enabled so the label = comma(count) works.
   geom_text(aes(label = comma(number_of_rides)), 
            vjust= .5,
            hjust = -.15, 
            family = "exo", 
            size = 4.5) + 
  labs(x = "Rider Status", 
       y = "Number of Rides", 
       title  = "Cyclistic Rides by Rider Status") + 
  # the following puts labels in thousands on y-axis (label = scales::comma)
  # and makes sure the graph starts on the axis. "limits =" sets the limit of the axis.
  scale_y_continuous(label = scales::comma, expand = c(0, 0), limits = c(0, 3500000))+  
  scale_fill_manual(values = c("indianred1", "goldenrod1"))+
  coord_flip() +
  theme_classic()+
  theme(text = element_text(family = "exo", size = 20),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(), 
        legend.position = "none")

length <- 
  ggplot(Cyclistic_CT, 
       aes(x = member_casual, 
           y = ride_time_avg, 
           fill = member_casual)) + 
  geom_bar(stat = "identity") +
  geom_text(aes(label = ride_time_avg), 
            vjust= .5,
            hjust = -.15,
            family = "exo",
            size = 4.5) +
  labs(x = "Rider Status", 
       y = "Average Length of Ride", 
       title  = "Average Length of Ride by Rider Status") + 
  scale_y_discrete(labels = NULL, breaks = NULL, expand = expansion(mult = c(0, .2)))+
  scale_fill_manual(values = c("indianred1", "goldenrod1"))+
  coord_flip() +
  theme_classic()+
  theme(legend.position = "none")+
  theme(text = element_text(family = "exo", size = 20))

number / length
```

## Most Popular Stations

Which stations are the most popular launching points for each type of rider?

```{r Most Popular Station by Rider Type, message=FALSE, warning=FALSE}

Pop_Station_Name <- Cyclistic %>% group_by(start_station_name, start_station_id, member_casual) %>% 
   summarize(count = n()) %>% select(start_station_name, start_station_id, member_casual, count) 
  
reactable(Pop_Station_Name, bordered = TRUE,
  highlight = TRUE, theme = fivethirtyeight(), searchable = TRUE, pageSizeOptions = 25, filterable = TRUE )
```

This table provides information that gives the starting station name, starting station id, whether the user was a member or a casual rider, and the number of rides started from that station based on whether the user was a member or casual rider. This is a very useful data frame from which we can draw important information.

For instance, we can identify the top ten starting stations for casual riders:

```{r Top 10 Starting Stations for Casual Riders}
casual_10 <- Pop_Station_Name %>% arrange(member_casual, desc(count))
casual_top_10 <- head(casual_10, 10)
table_casual_top_10 <- reactable(casual_top_10, bordered = TRUE,
  highlight = TRUE, theme = fivethirtyeight(), searchable = TRUE, filterable = TRUE)

chart_casual_top_10 <- ggplot(casual_top_10, aes(x = reorder(start_station_name, count), y = count, fill = count)) + 
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), color = "black", size = 6,
            vjust= .5,
            hjust = -.1, 
            family = "exo") +
  labs(y = "Number of Rides", fill = "Start Station Name",
       title  = "Most Popular Start Stations for Casual Riders") +
  scale_y_continuous(label = scales::comma, expand = c(0, 0), limits = c(0, 50000)) +
  coord_flip() +
  theme_classic()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(size = 15),
        axis.title.y = element_blank(),
        legend.position = "none", 
        text = element_text(family = "exo", size = 17)) +
  scale_fill_viridis(option = "magma", begin = .95, end = .5)

table_casual_top_10
chart_casual_top_10

```

And the top 10 starting stations for members:

```{r Top 10 Starting Station for Members}
member_10 <- Pop_Station_Name %>% arrange(desc(member_casual), desc(count))
member_top_10 <- head(member_10, 10)
table_member_top_10 <- reactable(member_top_10, bordered = TRUE,
  highlight = TRUE, theme = fivethirtyeight(), searchable = TRUE, filterable = TRUE)

chart_member_top_10 <- ggplot(member_top_10, aes(x = reorder(start_station_name, count), y = count, fill = count)) + 
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), color = "black", size = 6,
            vjust= .5,
            hjust = -.1, 
            family = "exo") +
  labs(y = "Number of Rides", fill = "Start Station Name",
            title  = "Most Popular Start Stations for Members") +
  # the following puts labels in thousands on y-axis (label = scales::comma)
  # and makes sure the graph starts on the axis. "limits =" sets the limit of the axis.
  scale_y_continuous(label = scales::comma, expand = c(0, 0), limits = c(0, 28000)) +
  coord_flip() +
  theme_classic()+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 15),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none", 
        text = element_text(family = "exo", size = 17))  +
  scale_fill_viridis(option = "magma", begin = .95, end = .5)

table_member_top_10
chart_member_top_10
```

If Cyclistic's goal is to increase membership, an advertising blitz in the areas in these top 10 lists would be advised. Of particular note is the number one spot for casual riders, Streeter Dr. & Grand Ave. It has twice as many trips as the next most popular spot, Millennium Park. Despite these eye-popping, and maybe suspicious, numbers, the data fits the context: the Streeter Dr. & Grand Ave. station is the station at Navy Pier, one of the most popular tourist destinations in Chicago.

The start station areas with high volumes of users might be areas ripe for marketing. So might the areas around start stations with high volumes of casual riders that also have long ride times. For casual riders, let's see what high volume start stations (over 500 rides) have the longest ride times.

```{r Casual Top 10 Start Stations by Ride Length, message=FALSE, warning=FALSE}
casual_avg_ride_length <- Cyclistic %>%  group_by(start_station_name, member_casual) %>% 
  filter(member_casual == "casual") %>% 
  summarise(count = n(), avg_ride_length = round(mean(ride_length_numeric), digits = 0)) %>% 
  mutate(sectime = seconds_to_period(avg_ride_length)) %>% 
  mutate(avg_ride_length_char = sprintf('%02d:%02d:%02d', sectime@hour, minute(sectime), second(sectime))) %>% 
  #sprintf is from stackoverflow - https://stackoverflow.com/questions/27312292/convert-seconds-to-days-hoursminutesseconds
  filter(count >= 500) %>% 
  select(start_station_name, count, member_casual, avg_ride_length_char) %>% 
  # ungroup() %>% 
  arrange(desc(avg_ride_length_char))

casual_avg_ride_length_top10 <-
  head(casual_avg_ride_length, 10)

reactable(casual_avg_ride_length_top10, bordered = TRUE,
  highlight = TRUE, theme = fivethirtyeight(), searchable = TRUE, filterable = TRUE)
```

Are there any stations that make an appearance in both the top 10 number of a casual rides and the top 10 average length of ride?

```{r Overlapping Top 10s}
casual_top_10s <- inner_join(casual_top_10, casual_avg_ride_length_top10, "start_station_name")
casual_top_10s
```

There is no overlapping station in the top 10 for casual riders by volume (number of rides) and the top 10 for casual riders by ride length. However, zoom out and include just the next one on the ride length list and you will find a very popular start station by volume: Millennium Park. A unique marketing strategy for Millennium Park might be beneficial, while Cyclistic might want to take different marketing angles towards the different stations on the respective lists above.

## Days of the Week

Let's now take a look at days of the week. Should Cyclistic focus advertising efforts on certain days of the week for casual riders? Are members more inclined to ride on weekdays or weekends?

```{r Rides by Day of Week, message=FALSE, warning=FALSE}
days_of_week <- Cyclistic %>% group_by(member_casual, weekday) %>% 
  summarise(count = n())
reactable(head(days_of_week, 14), bordered = TRUE,
  highlight = TRUE, theme = fivethirtyeight(), searchable = TRUE, filterable = TRUE, pagination = FALSE)
```

We can take a look at this and see...

```{r Rides by Day of Week Visualized, message=FALSE, warning=FALSE}
days_of_week %>% arrange(days_of_week) %>%  ggplot(aes(x= weekday, y = count, fill = member_casual)) +
  geom_histogram(stat = "identity") + 
  geom_text(aes(label = comma(count)), color = "white", size = 4, fontface = "bold",
                                                vjust= .5,
                                                hjust = 1.15, 
                                                family = "exo") +
  labs(y = "Number of Rides",
       title  = "Number of Rides by Day of Week") +
  # expand_limits(y=c(0, 50000)) +
  # put the discrete variables in the order you want them. 
  scale_x_discrete(limits = c("Saturday", "Friday", "Thursday", "Wednesday", "Tuesday", "Monday", "Sunday")) +
  coord_flip() +
  scale_fill_manual(values = c("indianred1", "goldenrod1")) +
  theme_classic()+
  theme(axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "none", 
        text = element_text(family = "exo", size = 20)) +
  facet_grid(. ~ member_casual)
```

...that casual riders are more likely to ride on the weekends. If targeting casual riders, advertising on the weekends is recommended.

## Time of Day by Day of Week

Another avenue to explore is time of day to target casual riders. We can accomplish this by taking a look at the time of day both members and casual riders are typically beginning their rides.

```{r Time of Day for Casual Riders and Members}
Cyclistic_hour_start <- Cyclistic  %>% 
transform (weekday = factor(weekday, levels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")))

Cyclistic_hour_start %>% ggplot(aes(x = hour, fill = member_casual)) +
  geom_bar()+
  labs(x = "Start Hour", y = "Number of Rides", title = "Number of Rides by Hour", 
       subtitle = "By Day") +
  scale_y_continuous(label = scales::comma, expand = c(0, 0), breaks = c(0, 30000, 60000), limits = c(0, 75000))+
  theme_classic() +
  scale_fill_manual(values = c("indianred1", "goldenrod1")) +
  theme(axis.ticks.y = element_blank(), 
        text = element_text(family = "exo"), 
        panel.spacing.y = unit(2, "lines")) +
  labs(fill = "Rider Type")+
  facet_wrap(~ weekday, scales = "free")
```

We see from these visualizations that people tend to sleep in on the weekends, but also causal rider traffic is most heavy in the middle of the day on Saturday and Sunday. On weekdays, casual riders traffic increases throughout the day, peaking in late afternoon or early evening. Advertising efforts should focus their efforts on these high traffic hour windows.

# Interactive Map on Tableau

Finally, [here is a link to an interactive dashboard on Tableau.](https://public.tableau.com/views/Top20StationswithMap/Top20withMap?:language=en-US&publish=yes&:display_count=n&:origin=viz_share_link) The top 20 start stations by total number of rides are given, along with the total hours ridden from each station. One can filter by rider type at the top. An interactive map is also included which shows number of rides by station (minimum of 100 rides). The rider type filter at the top also applies to the data presented in the map.

Useful information to be gleaned from the dashboard:

-   Not surprisingly, casual riders are practically non-existent in the months of January and February. Advertising directed at casual riders during this time is ill-advised.

-   There is a swell of casual riders during the summer months, particular around popular tourist destinations.

-   There is increased activity among members along Ellis Ave, south of the city center, during the fall months. The area of increased activity is in the vicinity of the University of Chicago medical complex. The increased rate could be tied to Covid cases and/or testing, but this information could be used to target casual riders as well, highlighting easy accessibility to health care facilities.

\*Because of imprecision in latitude and longitude data in the original data set, some ride totals may be slightly different than what is presented above. Some station names did not match the latitude and longitude coordinates, therefore there is some information missing, such as the Shedd Aquarium station from the top 20 stations for casual riders. Most of the same information from above was captured however, and the general conclusions remain the same.

# Moving Forward

The primary objective of the analysis was to determine behavioral differences between casual riders and members. After completing this analysis, the following behavioral differences were identified:

-   Casual riders have longer ride times on average than members

-   Many top 10 start stations by number of rides for casual riders are popular tourist destinations throughout city.

-   There is a disproportionate number of rides for casual riders on the weekend.

-   Most weekday rides for casual riders begin late afternoon to early evening. Most weekend rides begin midday through afternoon. This is in contrast to riders who are members who have higher number of rides during weekday mornings.

-   There is a higher volume of casual riders in the summer months.

## Deductions and Recommendations

It seems as if many casual riders are tourists, as gleaned from the volume of rides by time of year and popular start stations. It will be difficult to convince one-off weekend riders to sign up for a Cyclistic membership. Strategies to combat this challenge include:

-   Partnering with other business or services and presenting membership as part of a packaged deal.

-   Offering short-term membership options. If the only membership option is yearly, 3 month membership options might be more appealing to casual riders.

Marketing strategies have been suggested throughout the project, but one further strategy is dependent on how casual riders are currently charged.

-   If casual riders are charged by ride, advertising efforts should be directed toward high volume stations for casual riders - Streeter Dr. & Grand Ave., Millennium Park, Michigan Ave. & Oak St., etc. Highlight the benefit of membership for number of rides.

-   If casual riders are charged by length of ride, advertising efforts should be directed toward stations with long ride times - Museum of Science and Industry, 63rd St. Beach, etc. Highlight the benefit of membership for long ride times. If Cyclistic has an app, push notifications could be used to inform the rider how much money they would have saved on their long ride with a membership.

One possible avenue of further exploration is presence of other bike sharing ventures in other urban areas. A packaged deal for membership with other bike share ventures (similar to a previous suggestion) could entice casual riders to pursue membership. More data on bike share ventures in other urban areas would be needed for this analysis.
