---
title: "Descriptive_Statistics"
author: "Andrew Borozan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pacman)
p_load(tidyverse, skimr, scales, rmarkdown, magrittr, lubridate, janitor, ggrepel, ggthemes, showtext, knitr, reactable, palmerpenguins, car, broom, readxl, stringr, viridis, gghighlight)
```

Following along with a couple of different sources. Predominantly drawing from [statsandr.com](https://statsandr.com/blog/descriptive-statistics-in-r/).

Main points from the introduction:

-   Descriptive stats summarize and describe.

-   Identifying:

    -   Location measures - what is the central tendency? Around what point is the data gravitating?

    -   Dispersion measures - how spread out is the data? How far do the spiral arms go in this galaxy?

Looking at the iris data set:

```{r}
irisdat <- iris

irisdat
```

Start with min() and max().

```{r}
min(irisdat$Sepal.Length)

max(irisdat$Sepal.Length)
```

Or, range gives you both in one go:

```{r}
range(irisdat$Sepal.Length)
```

Mean and median:

```{r}
mean(irisdat$Sepal.Length, na.rm = TRUE, trim = .10)
# Won't compute if there are NAs. Also, get a trimmed mean with trim = .1. This will help with removing outliers. 

median(irisdat$Sepal.Length)

# Or with quantile function
quantile(irisdat$Sepal.Length, .5) 
# Can play with the quantile number to get what you want
quantile(irisdat$Sepal.Length, .9)

# First and third quartile
quantile(irisdat$Sepal.Length, .25)
quantile(irisdat$Sepal.Length, .75)

# And interquartile range
IQR(irisdat$Sepal.Length)
```

Standard deviation and variance:

```{r}
sd(irisdat$Sepal.Length)

var(irisdat$Sepal.Length)

# To compute the sd for all variables, use lapply()
lapply(irisdat[, 1:4], sd)
```

See it all at once with summary().

```{r}
summary(irisdat)

# And by group
irisdat %>% by(irisdat$Species, summary)
```

To see a whole slew of stats at once, use stat.desc() from pastecs.

```{r}
library(pastecs)
stat.desc(irisdat, norm = TRUE)
```

```{r}
irisdat$size <- ifelse(irisdat$Sepal.Length < median(irisdat$Sepal.Length),
  "small", "big"
)

table(irisdat$Species, irisdat$size)

round(prop.table(table(irisdat$Species, irisdat$size)), 2)
```

Mosaic plot - needs two qualitative variables:

```{r}
mosaicplot(table(irisdat$Species, irisdat$size), 
           color = TRUE)

library(vcd)
mosaic(~Species + size, 
       data = irisdat)
```

Scatterplot - two quantitative variables

```{r}
irisdat %>% 
  ggplot(aes(Sepal.Length, Petal.Length, color = Species)) +
  geom_point()
```

Use freq() and ctable() in summarytools package if you have at least one factor variable.

```{r out.width="100%"}
library(vcd)
Arthritis

Arthritis %>% count(Improved)

plot(Arthritis$Improved, horiz = TRUE)

Arthritis %>% count(Improved, Treatment)

Arthritis %>% group_by(Treatment, Improved) %>% 
  summarize(n = n()) %>%
  ggplot(aes(Treatment, n, fill = Improved)) +
  geom_col()

```

```{r}
qb <- read_csv(file.choose())
```

Correlation between two variables.

```{r}
irisdat %>% 
  ggplot(aes(Sepal.Length, Sepal.Width)) +
  geom_jitter(color = "blue") +
  geom_density2d(color = "white")


```

```{r}
irisdat %>% 
  ggplot(aes(Species, Sepal.Length, color = Species)) +
   geom_violin()+
  geom_jitter (width = .1)
 
```
